### 2023-07-25 article3.qmd
corrections following reviews

For the first time, BGC-Argo floats equipped with both the UVP6 and an OST were deployed in the Labrador Sea (LS) to investigate the seasonal and episodic dynamics of particles and carbon fluxes. The LS is a basin of the subpolar North Atlantic (NA) ocean which represents 20% of the global ocean carbon sink [@Takahashi2009-rv]. The efficiency of this sink is explained by both deep winter convection that traps CO$_{2}$ from the atmosphere [@Vage2008-qa; @Yashayaev2009-fe; @Yashayaev2017-mn] and by the largest spring phytoplankton bloom [@Siegel2002-hk] that consumes CO$_{2}$ for the photosynthetic production of organic matter. The triggering mechanism of the NA bloom was initially explained by the 'Critical Depth Hypothesis' of @Sverdrup1953-ki that defines a critical surface mixing depth where phytoplankton growth is equal to phytoplankton losses. According to @Sverdrup1953-ki, a bloom is therefore triggered when the surface mixed layer depth (MLD) is shallower than this critical depth resulting in a positive phytoplankton growth rate. However, this hypothesis was challenged by @Behrenfeld2010-my who suggested that the initiation of the NA bloom begins in winter with a slow phytoplankton biomass increase over the following months (i.e. 'Dilution-Recoupling' hypothesis). These hypotheses have been investigated with BGC-Argo floats but the debate still remains [@Mignot2018-gm; @Yang2020-hw; @Boss2010-do]. Distinct NA bloom phenologies were highlighted by @Lacour2015-da who used a climatology of satellite Chla to divide the NA into 6 bioregions. In their bioregionalization (see also their Figure 2), the LS is characterized by two bioregions with distinct bloom phenologies on both side of the 60°N parallel. The northern part is characterized by a short and potentially more intense spring bloom in late April whereas the bloom in the southern part starts in April-May and slowly develops to reach its maximum in June. @Lacour2015-da concluded that the concurrent influence of surface light and mixing was the first order mechanism controlling the onset of the spring bloom in both bioregions. This is coherent with the results of @Wu2008-vw and @Frajka-Williams2010-ph who found that the early bloom in the northern part is due to a very shallow mixed layer associated with low-salinity waters (i.e. haline stratification) while the southern bloom is related to the development of the seasonal thermocline (i.e. thermal stratification due to increased solar radiation). In late May 2022, two pairs of BGC-Argo floats were deployed in the southern part of the LS at the time of the spring bloom. While BGC-Argo floats typically drift at 1000 m for 10 days between each ascending profile, all floats were configured to drift at three distinct parking depths (200 m, 500 m and 1000 m) to study the downward POC flux and its attenuation derived from beam attenuance [@Estapa2023-oz] and to study particle size distributions (PSDs) and concentrations with the UVP6.

@@

All 4 floats have been configured to drift at 3 distinct parking depths (respectively, 200 m, 500 m and 1000 m for approximately 1, 3 and 5 days) to study the downward carbon flux and its attenuation. During these drift phases, measurements of particle abundance (14 size classes, between 0.1 mm and 2.5 mm) are taken every 20 minutes at 200 m and every 2h at 500 m and 1000 m whereas $c_{p}$ measurements are taken every 30 minutes at all depths. This configuration can be adjusted during the mission of the float via two-way Iridium communication. Finally, the vertical orientation of the transmissometer causes the accumulation of sinking material on the optical window. Therefore, a pump is cleaning the latter to remove the accumulated material between each drifting phase and subsequently avoid sensor drift.

@@

# read floats data
float_6904240 <- vroom('/home/flo/dox/ArgoShine/6904240_FromNetCDF.csv')
float_6904241 <- vroom('/home/flo/dox/ArgoShine/6904241_FromNetCDF.csv')
float_4903634 <- vroom('/home/flo/dox/ArgoShine/4903634_FromNetCDF.csv')
float_1902578 <- vroom('/home/flo/dox/ArgoShine/1902578_FromNetCDF.csv')

@@

p +
  geom_path(data = argo_map_geo, aes(lon, lat, colour = WMO), linewidth = 1) + 
  geom_point(data = argo_map_geo, aes(x= lon, y = lat, colour = WMO), size = 2) +
  scale_color_brewer(palette = 'Dark2') + theme_bw() +
  geom_point(data = transform_coord(argo_map_first), aes(lon, lat), shape = 17, size = 3) +
  geom_point(data = argo_map_last_geo, aes(lon, lat, colour = WMO), size = 4) +
  geom_point(data = argo_map_last_geo, aes(lon, lat), fill = 'black', size = 4, shape = 1) +
  #https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/#suppress-aesthetics-from-part-of-the-legend
  guides(color = guide_legend(override.aes = list(size = c(1.5,1.5,1.5,1.5),
                                                  shape = c(NA, NA, NA, NA)))) +
  annotation_scale(location = "br") + 
  annotation_north_arrow(location = "tr", which_north = "true") 
```
```{=latex}
\newpage
```

### Data processing

@@

Satellite Chla data were downloaded from a daily-merged cloud free GlobColour product (doi: 10.48670/moi-00281) with a spatial resolution of 4 km. MLD derived from satellite observations were downloaded from near-real time weekly data from the ARMOR3D product (doi: 10.48670/moi-00052) with a spatial resolution of 0.25°.

@@

up_part <- particles_bis |> filter(depth < 2000, size < 512) |>
  #group_by(size) |>
  ggplot() + facet_grid(wmo~size) + theme_bw() +
  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
  scale_y_reverse() +
  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
  labs(colour = '#/L', x = 'Month', y = 'Depth (m)')

@@

down_part <- particles_bis |> filter(depth < 2000, size >= 512) |>
  #group_by(size) |>
  ggplot() + facet_grid(wmo~size) + theme_bw() +
  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
  scale_y_reverse() +
  labs(colour = '#/L', x = 'Month', y = 'Depth (m)') +
  #geom_vline(xintercept = as_date('2022-01-01')) +
  scale_x_date(labels = date_format("%m"), date_breaks = '3 month')

@@

@@

p200m <- mean_uvp_data |> filter(park_depth == '200 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size, scales = 'free_y') #+
  #geom_vline(data = test, aes(xintercept = date_max))

@@

p200m <- mean_uvp_data |> filter(park_depth == '200 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size, scales = 'free_y') +
  geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')

@@

p500m <- mean_uvp_data |> filter(park_depth == '500 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size, scales = 'free_y') #+
  #geom_vline(data = test, aes(xintercept = date_max))

@@

p500m <- mean_uvp_data |> filter(park_depth == '500 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size, scales = 'free_y') +
  geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')

@@

p1000m <- mean_uvp_data |> filter(park_depth == '1000 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size, scales = 'free_y') #+
  #geom_vline(data = test, aes(xintercept = date_max))

@@

p1000m <- mean_uvp_data |> filter(park_depth == '1000 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size, scales = 'free_y') +
  geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')

@@

up <- tmp |> ggplot(aes(x = date, y = small_flux)) +
    geom_point(data = tmp, aes(x = date, y = small_flux, colour = WMO, shape = WMO)) +
  geom_smooth(data = tmp, aes(x = date, y = small_flux), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  guides(fill=guide_legend(title="WMO")) +
  facet_wrap(~park_depth, scales = 'free') +
  labs(x = '', y = TeX('$F_{small}$ (mg C m$^{-2}$ day$^{-1}$)')) +
  scale_y_continuous(trans = 'log10') +
  theme_bw()

down <- tmp |> ggplot(aes(x = date, y = large_flux)) +
  geom_point(data = tmp, aes(x = date, y = large_flux, colour = WMO, shape = WMO)) +
  geom_smooth(data = tmp, aes(x = date, y = large_flux), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  guides(fill=guide_legend(title="WMO")) +
  facet_wrap(~park_depth, scales = 'free') +
  labs(x = 'Month', y = TeX('$F_{large}$ (mg C m$^{-2}$ day$^{-1}$)')) +
  theme_bw() +
  scale_y_continuous(trans = 'log10') +
  theme(legend.position = 'none')

@@

# add MLD here
mld_data <- vroom('/home/flo/dox/ArgoShine/npq_corrected_data.csv')
# clean and smoothed chla data
mld_data <- mld_data |> select(cycle, MLD, pres, juld, wmo, chla_npq) |> dplyr::mutate(date = as_date(juld), WMO = factor(wmo, levels = c(6904240,4903634,6904241,1902578)))
mld_data <- mld_data |> group_by(date, WMO, pres, cycle) |> summarize(mean_chla = mean(chla_npq, na.rm=T), mld = unique(MLD))

@@

mld_plot <- ggplot(mld_data) + geom_line(aes(x = date, y = mld, colour = WMO), size = 1) +
  labs(x = 'Month', y = 'MLD (m)') + scale_y_reverse() +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + theme_bw() +
  geom_hline(yintercept = 200, colour = 'black', linetype = 'dashed') +
  geom_hline(yintercept = 500, colour = 'black', linetype = 'dashed')

@@

@@

chla_plot_mld <- ggplot(chla_mld) + geom_line(aes(x = date, y = chla_0_mld, colour = as.factor(wmo)), linewidth = 1) + labs(x = 'Month', y = TeX('Integrated Chla content (mg m$^{-2}$)'), colour = 'WMO') +
  theme_bw() + scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  #scale_colour_manual(values = c('#003366', '#E31B23'))
  scale_color_brewer(palette = 'Dark2')

chla_plot_100m <- ggplot(chla_mld) + geom_line(aes(x = date, y = chla_0_100m, colour = as.factor(wmo)), linewidth = 1) + labs(x = 'Month', y = TeX('Integrated Chla content (mg m$^{-2}$)'), colour = 'WMO') +
  theme_bw() + scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  scale_color_brewer(palette = 'Dark2') + theme(legend.position = 'none')

@@

Still, it is unclear why we observe a progressive emptying of the water column starting below 500 m for size classes below 323 $\mu$m (@fig-uvp6-profiling-102-406). This pattern could result from several causes. It could either be the signature of a subduction event [@Llort2018-pa] either the end of a slow sinking export (i.e latent dynamics of the system) or a sign of fragmentation processes by zooplankton (i.e. zooplankton feeding). Therefore, we looked at physical properties (density anomaly and spiciness) and well as particle characteristics (zooplankton concentration and spectral slope) to investigate this particularity (@fig-spice). Spiciness is a variable that allows the differentiation of water masses with distinct thermohaline properties but similar density [@Flament2002-zm]. @fig-spice shows that both the density and the spiciness fields are uniform below 500 m hence it is unlikely that a subduction event would be the reason for this decrease in particle abundance below 500 m. Using the embedded zooplankton classification results from the UVP6, we looked at concentrations of *Calanoida*, a species that is abundant in the NA [@Corkett1979-ha]. @fig-spice shows that the concentration of *Calanoida* is higher from the end of May to July in the surface layer. There also seems to be a lower patch between 250 and 500 m from August to November however it is difficult to estimate their impact on particle concentrations at this stage. Finally, we computed the spectral slope to look for aggregation/fragmentation processes. The spectral slope defines the slope of the PSD where a high (in absolute values) spectral slope indicates that small particules dominate the PSD. Here, the panel on spectral slopes in @fig-spice shows a lower (in absolute values) spectral slope at the beginning of the time series but it is also difficult to relate it to biological processes such as zooplankton feeding without any additional data.

>>>>

For the first time, BGC-Argo floats equipped with both the UVP6 and an OST were deployed in the Labrador Sea (LS) to investigate the seasonal and episodic dynamics of particles and carbon fluxes. The LS is a basin of the subpolar North Atlantic (NA) ocean which represents 20% of the global ocean carbon sink [@Takahashi2009-rv]. The efficiency of this sink is explained by both deep winter convection that traps CO$_{2}$ from the atmosphere [@Vage2008-qa; @Yashayaev2009-fe; @Yashayaev2017-mn] and by the largest spring phytoplankton bloom [@Siegel2002-hk] that consumes CO$_{2}$ for the photosynthetic production of organic matter. The triggering mechanism of the NA bloom was initially explained by the 'Critical Depth Hypothesis' of @Sverdrup1953-ki that defines a critical surface mixing depth where phytoplankton growth is equal to phytoplankton losses. According to @Sverdrup1953-ki, a bloom is therefore triggered when the surface mixed layer depth (MLD) is shallower than this critical depth resulting in a positive phytoplankton growth rate. However, this hypothesis was challenged by @Behrenfeld2010-my who suggested that the initiation of the NA bloom begins in winter with a slow phytoplankton biomass increase over the following months (i.e. 'Dilution-Recoupling' hypothesis). These hypotheses have been investigated with BGC-Argo floats but the debate still remains [@Mignot2018-gm; @Yang2020-hw; @Boss2010-do]. Distinct NA bloom phenologies were highlighted by @Lacour2015-da who used a climatology of satellite Chla to divide the NA into 6 bioregions. In their bioregionalization (see also their Figure 2), the LS is characterized by two bioregions with distinct bloom phenologies on both side of the 60°N parallel. The northern part is characterized by a short and potentially more intense spring bloom in late April whereas the bloom in the southern part starts in April-May and slowly develops to reach its maximum in June. @Lacour2015-da concluded that the concurrent influence of surface light and mixing was the first order mechanism controlling the onset of the spring bloom in both bioregions. This is coherent with the results of @Wu2008-vw and @Frajka-Williams2010-ph who found that the early bloom in the northern part is due to a very shallow mixed layer associated with low-salinity waters (i.e. haline stratification) while the southern bloom is related to the development of the seasonal thermocline (i.e. thermal stratification due to increased solar radiation). In late May 2022, two pairs of BGC-Argo floats were deployed in the southern part of the LS at the time of the spring bloom. While BGC-Argo floats typically drift at 1000 m for 10 days between each ascending profile, all floats were configured to drift at three distinct parking depths (200 m, 500 m and 1000 m) to study the downward POC flux and its attenuation derived from beam attenuance [@Estapa2013-gw; @Estapa2019-nk] and to study particle size distributions (PSDs) and concentrations with the UVP6.

@@

All 4 floats have been configured to drift at 3 distinct parking depths (respectively, 200 m, 500 m and 1000 m for approximately 1, 3 and 5 days) to study the downward carbon flux and its attenuation. During these drift phases, measurements of particle abundance (14 size classes, between 0.1 mm and 2.5 mm) are taken every 20 minutes at 200 m and every 2h at 500 m and 1000 m whereas $c_{p}$ measurements are taken every 30 minutes at all depths. This configuration can be adjusted during the mission of the float via two-way Iridium communication. Finally, the vertical orientation of the transmissometer causes the accumulation of sinking material on the optical window. Therefore, a dedicated pump is used to clean the latter to remove the accumulated material between each drifting phase and subsequently avoid sensor drift.

@@

# read floats data
float_6904240 <- vroom('/home/flo/dox/ArgoShine/SAVE_FOR_MANUSCRIPT/6904240_FromNetCDF.csv')
float_6904241 <- vroom('/home/flo/dox/ArgoShine/SAVE_FOR_MANUSCRIPT/6904241_FromNetCDF.csv')
float_4903634 <- vroom('/home/flo/dox/ArgoShine/SAVE_FOR_MANUSCRIPT/4903634_FromNetCDF.csv')
float_1902578 <- vroom('/home/flo/dox/ArgoShine/SAVE_FOR_MANUSCRIPT/1902578_FromNetCDF.csv')

# float_6904240 <- vroom('/home/flo/dox/ArgoShine/6904240_FromNetCDF.csv')
# float_6904241 <- vroom('/home/flo/dox/ArgoShine/6904241_FromNetCDF.csv')
# float_4903634 <- vroom('/home/flo/dox/ArgoShine/4903634_FromNetCDF.csv')
# float_1902578 <- vroom('/home/flo/dox/ArgoShine/1902578_FromNetCDF.csv')

@@

p +
  geom_path(data = argo_map_geo, aes(lon, lat, colour = WMO), linewidth = 1) + 
  geom_point(data = argo_map_geo, aes(x= lon, y = lat, colour = WMO), size = 2) +
  #scale_color_brewer(palette = 'Dark2') + 
  theme_bw() +
  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77')) +
  geom_point(data = transform_coord(argo_map_first), aes(lon, lat), shape = 17, size = 3) +
  geom_point(data = argo_map_last_geo, aes(lon, lat, colour = WMO), size = 4) +
  geom_point(data = argo_map_last_geo, aes(lon, lat), fill = 'black', size = 4, shape = 1) +
  #https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/#suppress-aesthetics-from-part-of-the-legend
  guides(color = guide_legend(override.aes = list(size = c(1.5,1.5,1.5,1.5),
                                                  shape = c(NA, NA, NA, NA)))) +
  annotation_scale(location = "br") + 
  annotation_north_arrow(location = "tr", which_north = "true") 
```

```{=latex}
\newpage
```
### Data processing

@@

Satellite Chla data were downloaded from a daily-merged cloud free GlobColour product (doi: 10.48670/moi-00281) with a spatial resolution of 4 km. MLD data were downloaded from near-real time weekly data from the ARMOR3D product (doi: 10.48670/moi-00052) with a spatial resolution of 0.25°.

@@

particles_bis <- particles_bis |> mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)))

facet_names_up <- c(`102` = '102 - 128 µm',
                 `128` = '128 - 161 µm',
                 `161` = '161 - 203 µm',
                 `203` = '203 - 256 µm',
                 `256` = '256 - 323 µm',
                 `323` = '323 - 406 µm',
                 `406` = '406 - 512 µm',
                 `6904240` = '6904240',
                 `6904241` = '6904241',
                 `4903634` = '4903634',
                 `1902578` = '1902578')

up_part <- particles_bis |> filter(depth < 2000, size < 512) |>
  #group_by(size) |>
  ggplot() + facet_grid(wmo~size, labeller = as_labeller(facet_names_up)) + theme_bw() +
  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
  scale_y_reverse() +
  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
  labs(colour = '#/L', x = 'Month', y = 'Depth (m)')

@@

facet_names_down <- c(`512` = '512 - 645 µm',
                 `645` = '645 - 813 µm',
                 `813` = '0.81 - 1.02 mm',
                 `1020` = '1.02 - 1.29 mm',
                 `1290` = '1.29 - 1.63 mm',
                 `1630` = '1.63 - 2.05 mm',
                 `2050` = '2.05 - 2.50 mm',
                 `6904240` = '6904240',
                 `6904241` = '6904241',
                 `4903634` = '4903634',
                 `1902578` = '1902578')

down_part <- particles_bis |> filter(depth < 2000, size >= 512) |>
  #group_by(size) |>
  ggplot() + facet_grid(wmo~size, labeller = as_labeller(facet_names_down)) + theme_bw() +
  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + 
  scale_y_reverse() +
  labs(colour = '#/L', x = 'Month', y = 'Depth (m)') +
  #geom_vline(xintercept = as_date('2022-01-01')) +
  scale_x_date(labels = date_format("%m"), date_breaks = '3 month')

@@

facet_all <- c(`102` = '102 - 128 µm',
                 `128` = '128 - 161 µm',
                 `161` = '161 - 203 µm',
                 `203` = '203 - 256 µm',
                 `256` = '256 - 323 µm',
                 `323` = '323 - 406 µm',
                 `406` = '406 - 512 µm',
                 `512` = '512 - 645 µm',
                 `645` = '645 - 813 µm',
                 `813` = '0.81 - 1.02 mm',
                 `1020` = '1.02 - 1.29 mm',
                 `1290` = '1.29 - 1.63 mm',
                 `1630` = '1.63 - 2.05 mm',
                 `2050` = '2.05 - 2.50 mm')

@@

p200m <- mean_uvp_data |> filter(park_depth == '200 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  #scale_color_brewer(palette = 'Dark2') + 
  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77')) +
  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size, scales = 'free_y', labeller = as_labeller(facet_all)) #+
  #geom_vline(data = test, aes(xintercept = date_max))

@@

p200m <- mean_uvp_data |> filter(park_depth == '200 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  #scale_color_brewer(palette = 'Dark2') + 
  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77')) +
  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size, scales = 'free_y', labeller = as_labeller(facet_all)) +
  geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')

@@

p500m <- mean_uvp_data |> filter(park_depth == '500 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  #scale_color_brewer(palette = 'Dark2') + 
  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77')) +
  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size, scales = 'free_y', labeller = as_labeller(facet_all)) 
  #geom_vline(data = test, aes(xintercept = date_max))

@@

p500m <- mean_uvp_data |> filter(park_depth == '500 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  #scale_color_brewer(palette = 'Dark2') + 
  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77')) +
  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size, scales = 'free_y', labeller = as_labeller(facet_all)) +
  geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')

@@

p1000m <- mean_uvp_data |> filter(park_depth == '1000 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  #scale_color_brewer(palette = 'Dark2') + 
  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77')) +
  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size, scales = 'free_y', labeller = as_labeller(facet_all)) #+
  #geom_vline(data = test, aes(xintercept = date_max))

@@

p1000m <- mean_uvp_data |> filter(park_depth == '1000 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  #scale_color_brewer(palette = 'Dark2') + 
  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77')) +
  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
  theme_bw() + labs(x = 'Month', y = 'Particle abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size, scales = 'free_y', labeller = as_labeller(facet_all)) +
  geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')

@@

up <- tmp |> ggplot(aes(x = date, y = small_flux)) +
    geom_point(data = tmp, aes(x = date, y = small_flux, colour = WMO, shape = WMO)) +
  geom_smooth(data = tmp, aes(x = date, y = small_flux), colour = 'black') +
  #scale_color_brewer(palette = 'Dark2') + 
  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77')) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  guides(fill=guide_legend(title="WMO")) +
  facet_wrap(~park_depth, scales = 'free') +
  labs(x = '', y = TeX('$F_{small}$ (mg C m$^{-2}$ day$^{-1}$)')) +
  scale_y_continuous(trans = 'log10') +
  theme_bw()

down <- tmp |> ggplot(aes(x = date, y = large_flux)) +
  geom_point(data = tmp, aes(x = date, y = large_flux, colour = WMO, shape = WMO)) +
  geom_smooth(data = tmp, aes(x = date, y = large_flux), colour = 'black') +
  #scale_color_brewer(palette = 'Dark2') + 
  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77')) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  guides(fill=guide_legend(title="WMO")) +
  facet_wrap(~park_depth, scales = 'free') +
  labs(x = 'Month', y = TeX('$F_{large}$ (mg C m$^{-2}$ day$^{-1}$)')) +
  theme_bw() +
  scale_y_continuous(trans = 'log10') +
  theme(legend.position = 'none')

@@

# add MLD here
mld_data <- vroom('/home/flo/dox/ArgoShine/npq_corrected_data.csv')
#filter date until March
mld_data <- mld_data[mld_data$juld < "2023-03-05",]
# clean and smoothed chla data
mld_data <- mld_data |> select(cycle, MLD, pres, juld, wmo, chla_npq) |> dplyr::mutate(date = as_date(juld), WMO = factor(wmo, levels = c(6904240,4903634,6904241,1902578)))
mld_data <- mld_data |> group_by(date, WMO, pres, cycle) |> summarize(mean_chla = mean(chla_npq, na.rm=T), mld = unique(MLD))

@@

mld_plot <- ggplot(mld_data) + geom_line(aes(x = date, y = mld, colour = WMO), size = 1) +
  labs(x = 'Month', y = 'MLD (m)') + scale_y_reverse() +
  #scale_color_brewer(palette = 'Dark2') + 
  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77')) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + theme_bw() +
  geom_hline(yintercept = 200, colour = 'black', linetype = 'dashed') +
  geom_hline(yintercept = 500, colour = 'black', linetype = 'dashed')

@@

#filter date until March
npq_data <- npq_data[npq_data$juld < "2023-03-05",]

@@

chla_plot_mld <- ggplot(chla_mld) + geom_line(aes(x = date, y = chla_0_mld, colour = as.factor(wmo)), linewidth = 1) + labs(x = 'Month', y = TeX('Integrated Chla content (mg m$^{-2}$)'), colour = 'WMO') +
  theme_bw() + scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  #scale_colour_manual(values = c('#003366', '#E31B23'))
  #scale_color_brewer(palette = 'Dark2') 
  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77'))

chla_plot_100m <- ggplot(chla_mld) + geom_line(aes(x = date, y = chla_0_100m, colour = as.factor(wmo)), linewidth = 1) + labs(x = 'Month', y = TeX('Integrated Chla content (mg m$^{-2}$)'), colour = 'WMO') +
  theme_bw() + scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  #scale_color_brewer(palette = 'Dark2') + 
  scale_colour_manual(values = c('#D95F02', '#7570B3', '#E6AB02', '#1B9E77')) +
  theme(legend.position = 'none')

@@

Still, it is unclear why we observe a progressive emptying of the water column starting below 500 m for size classes below 323 $\mu$m (@fig-uvp6-profiling-102-406). This pattern could result from several causes. It could either be the signature of a subduction event [@Llort2018-pa] either the end of a slow sinking export (i.e seasonal dynamics of the system) or a sign of fragmentation processes by zooplankton (i.e. zooplankton feeding). Therefore, we looked at physical properties (density anomaly and spiciness) and well as particle characteristics (zooplankton concentration and spectral slope) to investigate this particularity (@fig-spice). Spiciness is a variable that allows the differentiation of water masses with distinct thermohaline properties but similar density [@Flament2002-zm]. @fig-spice shows that both the density and the spiciness fields are uniform below 500 m hence it is unlikely that a subduction event would be the reason for this decrease in particle abundance below 500 m. Using the embedded zooplankton classification results from the UVP6, we looked at concentrations of *Calanoida*, a species that is abundant in the NA [@Corkett1979-ha]. @fig-spice shows that the concentration of *Calanoida* is higher from the end of May to July in the surface layer. There also seems to be a lower patch between 250 and 500 m from August to November however it is difficult to estimate their impact on particle concentrations at this stage because more data are needed. Finally, we computed the spectral slope to look for aggregation/fragmentation processes. The spectral slope defines the slope of the PSD where a high (in absolute values) spectral slope indicates that small particules dominate the PSD. Here, the panel on spectral slopes in @fig-spice shows a lower (in absolute values) spectral slope (i.e. large particles dominate) at the beginning of the time series but it is also difficult to relate it to biological processes such as zooplankton feeding without any additional data.
